---
title: "map creation"
author: "Alina Kereszt"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include=FALSE}
# install packages
# options("install.lock"=FALSE)
# install.packages("tidyverse")
# install.packages("leaflet")
# install.packages("geosphere")
# install.packages("measurements")
```

```{r, include=FALSE}
# load packages
library(tidyverse)
library(leaflet)
library(geosphere)
library(RColorBrewer)
library(measurements)
```

# Read data
These data files are the result of 'cleaning.rmd' and 'distance_calc.rmd'. Please go through those files if you have not yet done so.
```{r}
flats <- read.csv("data/flats_w_distances.csv")
stops <- read.csv("data/stops_revised.csv")
```

# Parameter exploration
**In this section, you can see the optional parameters. In order to run the selection function, please adjust the contents of each selection to suit what you are looking for.**

## Overview of whole dataset
```{r}
str(flats)
```

## Options selection
Always keep the values that apply. E.g. if you don't care whether a flat allows animals, you can keep c("yes", "only small pets", "no", NA), however if you have a medium-sized dog, you would want to keep only c("yes"). Please try not to add contradicting information, e.g. a 12 m2 flat with 5 rooms, or a flat in the 23rd quarter that is 500 m from the center, as you will receive an error later on.

Here, the parameters are set to base values:
```{r}
# Does the flat allow pets?
demo_animals_choice <- c("yes", 
                    "only small pets", 
                    "no", 
                    NA)

# What is the condition of the flat?
demo_condition_choice <- c("newly built",
                      "like new",
                      "renovated",
                      "good",
                      "in need of renovation",
                      NA)

# Does the building have an elevator?
demo_elevator_choice <- c("yes",
                     "no",
                     NA)

# Which quarter of Budapest is the flat located in?
demo_quarter_choice <- c(1, 2, 3, 4, 5,
                    6, 7, 8, 9, 10,
                    11, 12, 13, 14, 15,
                    16, 17, 18, 19, 20,
                    21, 22, 23)

# What is the price range for the apartment?
demo_min_price <- 50000
demo_max_price <- 2000000

# How many rooms does the flat have?
demo_rooms_choice <- c(1, 2, 3, 4, 5)

# How big is the flat?
demo_min_size <- 12
demo_max_size <- 500

# How far in meters is the flat from the center of Budapest?
demo_min_distance_center <- 0
demo_max_distance_center <- 18000

# How far in meters is the flat from the nearest public transport stop of any kind?
demo_min_distance_stop_any <- 0
demo_max_distance_stop_any <- 700

# How far in meters is the flat from the nearest railway-type public transport stop?
demo_min_distance_stop_rail <- 0
demo_max_distance_stop_rail <- 7000
```

Here, an example is shown. In this case, the user has the following stipulations:

- able to keep small pets
- the building is in some kind of good condition
- the building has an elevator
- price not higher than 150 000 HUF/month
- at least 2 rooms
- at least 40 m2 but no greater than 120 m2
- at most 5 km from the center
- at most 100 m from the nearest public transport stop of any kind
```{r}
# Does the flat allow pets?
my_animals_choice <- c("yes", 
                    "only small pets")

# What is the condition of the flat?
my_condition_choice <- c("newly built",
                      "like new",
                      "renovated",
                      "good")

# Does the building have an elevator?
my_elevator_choice <- c("yes")

# Which quarter of Budapest is the flat located in?
my_quarter_choice <- c(1, 2, 3, 4, 5,
                    6, 7, 8, 9, 10,
                    11, 12, 13, 14, 15,
                    16, 17, 18, 19, 20,
                    21, 22, 23)

# What is the price range for the apartment?
my_min_price <- 50000
my_max_price <- 150000

# How many rooms does the flat have?
my_rooms_choice <- c(2, 3, 4, 5)

# How big is the flat?
my_min_size <- 40
my_max_size <- 120

# How far in meters is the flat from the center of Budapest?
my_min_distance_center <- 0
my_max_distance_center <- 5000

# How far in meters is the flat from the nearest public transport stop of any kind?
my_min_distance_stop_any <- 0
my_max_distance_stop_any <- 100

# How far in meters is the flat from the nearest railway-type public transport stop?
my_min_distance_stop_rail <- 0
my_max_distance_stop_rail <- 7000
```

# Define function
```{r}
show_options <- function(data,
                        stops_data,
                        animals_choice,
                        condition_choice,
                        elevator_choice,
                        quarter_choice,
                        min_price,
                        max_price,
                        rooms_choice,
                        min_size,
                        max_size,
                        min_distance_center,
                        max_distance_center,
                        min_distance_stop_any,
                        max_distance_stop_any,
                        min_distance_stop_rail,
                        max_distance_stop_rail) {
  
  ##### FILTER DATA BY INPUT PARAMETERS
  selection <- flats %>% 
  
    # apply animals filter
    filter(animals %in% animals_choice) %>% 
    
    # apply condition filter
    filter(condition %in% condition_choice) %>% 
    
    # apply elevator choice
    filter(elevator %in% elevator_choice) %>% 
    
    # apply quarter choice
    filter(quarter %in% quarter_choice) %>% 
    
    # apply price choice
    filter(price >= min_price & price <= max_price) %>% 
    
    # apply room number choice
    filter(rooms %in% rooms_choice) %>% 
    
    # apply size choice
    filter(size >= min_size & size <= max_size) %>% 
    
    # apply distance from center choice
    filter(distance_to_bud_center >= min_distance_center & distance_to_bud_center <= max_distance_center) %>% 
    
    # apply distance from any public transport filter
    filter(nearest_stop_distance >= min_distance_stop_any & nearest_stop_distance <= max_distance_stop_any) %>% 
    
    # apply distance from rails transport filter
    filter(nearest_stop_distance_rail >= min_distance_stop_rail & nearest_stop_distance_rail <= max_distance_stop_rail)
  
  ##### PLOT
  if (nrow(selection) == 0) {
    
    print("There are no listings that match your criteria. Please broaden your search parameters.")
    
  } else {
    
    mymap <- leaflet() %>%
      addTiles() %>% 
      setView(
        lng = 19.0513, 
        lat = 47.4925,
        zoom = 12)
    
    for (i in seq_len(nrow(selection))) {
      lat <- selection$lat[i]
      lon <- selection$lon[i]
      url <- selection$url[i]
      price <- selection$price[i]
      
      mymap <- mymap %>%
        addMarkers(lng = lon, 
                   lat = lat,
                   popup = paste0(price,
                                  " HUF, ",
                                  "<a href='", 
                                  url, 
                                  "' target='_blank'>See listing</a>"))
    }
    
    for (i in seq_len(nrow(stops))) {
      lat <- stops$lat[i]
      lon <- stops$lon[i]
      routes <- stops$routes[i]
      
      mymap <- mymap %>%
        addCircleMarkers(lng = lon, 
                         lat = lat,
                         radius = routes * 0.3, 
                         fillOpacity = 0.3,
                         color = "blue", 
                         stroke = FALSE)
    }
    
    mymap
    
  }
}
```

# Run function with self-set parameters
```{r}
show_options(flats,
             stops,
             my_animals_choice,
             my_condition_choice,
             my_elevator_choice,
             my_quarter_choice,
             my_min_price,
             my_max_price,
             my_rooms_choice,
             my_min_size,
             my_max_size,
             my_min_distance_center,
             my_max_distance_center,
             my_min_distance_stop_any,
             my_max_distance_stop_any,
             my_min_distance_stop_rail,
             my_max_distance_stop_rail)
```























